### 2022-06-21 preprint.tex
Deploying to gh-pages from @ PoisotLab/ms_betacov_mapping@1b4ddd7d8db0b6e031f2dc59c93240bedff693b2 🚀
Deploying to gh-pages from @ PoisotLab/ms_betacov_mapping@16e453a1bff6288438ff77acaba93a8c85556ae7 🚀

\noindent {\LARGE \textbf{\textsf{The coevolutionary mosaic of
\emph{Betacoronavirus} spillover risk from bat hosts}}}

@@

We downloaded the data on bats hosts of \emph{Betacoronavirus} assembled
by Becker et al. (2022) from
\texttt{https://www.viralemergence.org/betacov} on Apr.~2022, and
filtered it to ``known'' hosts (established before the emergence of
SARS-CoV-2) and ``novel'' hosts (confirmed through sampling and
competence assays since the initial data collection). The original
database was assembled by a combination of data mining and literature
surveys, including automated alerts on the ``bats'' and ``coronavirus''
keywords to identify novel empirical evidence of bats-betacoronaviruses
associations; this yielded a total of 126 known hosts, 47 of which were
novel hosts.

@@

We downloaded the rangemap of every current bat species that was
classified as an empirically documented host of \emph{Betacoronavirus}
from the previous step, according to recent IUCN data (IUCN 2021). The
range maps were subsequently rasterized using the \texttt{rasterize}
function from \texttt{GDAL} (Rouault et al. 2022) at a resolution of
approximately 100kmx100km. For every pixel in the resulting raster where
at least one bat host of \emph{Betacoronavirus} was present, we extract
the species pool (list of all known bat hosts), which was used to
calculate the following risk assessment components: bat phylogenetic
diversity, bat compositional uniqueness, and predicted viral sharing
risk.

@@

For every pixel, we measured Faith's Phylogenetic Diversity (Faith 1992)
based on a recent synthetic tree with robust time calibration, covering
about 6000 mammalian species (Upham et al. 2019). Faith's PD measures
the sum of unique branches from an arbitrary root to a set of tips, and
comparatively larger values indicate a more phylogenetic diverse species
pool. We measured phylogenetic diversity starting from the root of the
entire tree (and not from Chiroptera); this bears no consequences on the
resulting values, since all branches leading up to Chiroptera are only
counted one per species pool, and (as we explain when describing the
assembly of the composite risk map), all individual risk components are
ranged in {[}0,1{]}. This measure incorporates a richness component,
which we chose not to correct for; the interpretation of the
phylogenetic diversity is therefore a weighted species richness, that
accounts for phylogenetic over/under-dispersal in some places.

@@

For every species pool, we measured its Local Contribution to
Beta-Diversity (Legendre and De Cáceres 2013); LCBD works from a
species-data matrix (traditionally noted as \(\mathbf{Y}\)), where
species are rows and sites are columns, and a value of 1 indicates
occurrence. We extracted the Y matrix assuming that every pixel
represents a unique location, and following best practices (Legendre and
Condit 2019) transformed it using Hellinger's distance to account for
unequal bat richness at different pixels. The correction of raw
community data is particularly important for two reasons: first, it
prevents the artifact of richer sites having higher importance; second,
it removes the effect of overall species richness, which is already
incorporated in the phylogenetic diversity component. High values of
LCBD indicate that the pixel has a community that is on average more
dissimilar in species composition than what is expected knowing the
entire matrix, i.e.~a more unique community. Recent results by Dansereau
et al. (2022) shows that LCBD measures are robust with regards to
spatial scale, and are therefore applicable at the global scale.

@@

For all bat hosts of \emph{Betacoronavirus}, we extracted their
predicted viral sharing network, generated from a previously published
generalized additive mixed model of virus sharing by a tensor function
of phylogenetic distance and geographic range overlap (Albery et al.
2020). This network stores pairwise values of viral community
similarity. To project viral sharing values into a single value for
every pixel, we averaged the pairwise scores. High values of the average
sharing propensity means that this specific extant bat assemblage is
likely to be proficient at exchanging viruses.

@@

To visualize the aggregated risk at the global scale, we combine the
three individual risk components (phylogenetic diversity, compositional
uniqueness, and viral sharing) using an additive color model (Seekell et
al. 2018). In this approach, every risk component gets assigned a
component in the RGB color model (phylogenetic diversity is green,
compositional uniqueness is red, and viral sharing is blue). In order to
achieve a valid RGB measure, all components are re-scaled to the
{[}0,1{]} interval, so that a pixel with no sharing, no phylogenetic
diversity, and no compositional uniqueness is black, and a pixel with
maximal values for each is white. This additive model conveys both the
intensity of the overall risk, but also the nature of the risk as colors
diverge towards combinations of values for three risk components. Out of
the possible combinations, the most risky in terms or rapid
diversification and spillover potential is high phylogenetic diversity
and low viral sharing (Gomulkiewicz et al. 2000, Cavender-Bares et al.
2009), in that this allows multiple independent host-virus
coevolutionary dynamics to take place in the same location. In the
colorimetric space, this correspond to yellow -- because the HSV space
is more amenable to calculations for feature extraction (see \emph{e.g.}
Keke et al. 2010), we measured the risk level by calculating the angular
distance of the hue of each pixel to a reference value of 60, and
weighted this risk level by the value component. Specifically, given a
pixel with colorimetric coordinates \((h,s,v)\), its ranged weighted
risk value is

@@

To next represent phylogeography of betacoronaviruses in bats, we
aggregated and analyzed betacoronavirus sequence data. We used the
following query to pull all \emph{Betacoronavirus} sequence data from
the GenBank Nucleotide database except SARS-CoV-2;
(``Betacoronavirus''{[}Organism{]} OR betacoronavirus{[}All Fields{]})
NOT (``Severe acute respiratory syndrome coronavirus 2''{[}Organism{]}
OR sars-cov-2{[}All Fields{]}). We added a single representative
sequence for SARS-CoV-2 and manually curated to remove sequences without
the RNA-dependent RNA polymerase (RdRp) sequence or that contained words
indicating recombinant or laboratory strains including ``patent,''
``mutant,'' ``GFP,'' and ``recombinant.'' We filtered over-represented
taxa including betacoronavirus 1, hCoV-OC43, Middle East respiratory
syndrome coronavirus, Murine hepatitis virus, and hCoV-HKU1. Curated
betacoronavirus RdRp sequences were then aligned using MAFFT v 1.4.0
{[}Katoh and Standley (2013); Algorithm FFT-NS-2, Scoring matrix 200PAM
/ k=2, gap open penalty 1.53m offset value 0.123{]} and a maximum
likelihood tree reconstructed in IQ-TREE v 1.6.12 (Nguyen et al. 2015)
with ModelFinder (Kalyaanamoorthy et al. 2017) ultrafast bootstrap
approximation (Hoang et al. 2018) with a general time reversible model
with empirical base frequencies and the 5-discrete-rate-category
FreeRaye model of nucleotide substitution (GTR+F+R5).

We first tested the hypothesis that hotspots of viral diversification
would track hotspots of bat diversification. To do so, we plotted the
number of known bat hosts (specifically only those included in the
phylogeny, so there was a 1:1 correspondence between data sources)
against the ``mean evolutionary distinctiveness'' of the associated
viruses. To calculate this, we derived the fair proportions evolutionary
distinctiveness (Isaac et al. 2007) for each of the viruses in the tree,
then averaged these at the bat species level, projected these values
onto their geographic distributions, and averaged across every bat found
in a given pixel. As such, this can be thought of as a map of the mean
evolutionary distinctiveness of the known viral community believed to be
associated with a particular subset of bats present.

@@

Subsequently, we tested the hypothesis that the biogeography of bat
betacoronaviruses should track the biogeography of their hosts. To test
this idea, we loosely adapted a method from (Kreft and Jetz 2007, 2010),
who proposed a phylogenetic method for the delineation of animal
biogeographic regions. In their original method, a distance matrix -
where each row or column represents a geographic raster's grid cell, and
the dissimilarity values are the ``beta diversity similarity'' of their
community assemble - undergoes non-metric multidimensional scaling
(NMDS); the first two axes of the NMDS are projected geographically
using a four-color bivariate map. Here, we build on this idea with an
entirely novel methodology. First, we measure the phylogenetic distance
between the different viruses in the betacoronaviruses tree by using the
cophenetic function in \texttt{ape} (Paradis and Schliep 2019);
subsequently, we take a principal components analysis of that distance
matrix (readily interchangeable for NMDS in this case) to project the
viral tree into an n-dimensional space. We then take the first two
principal components and, as with the evolutionary distinctiveness
analysis, aggregated these to a mean host value and projected them using
a four-color bivariate map.

@@

\hypertarget{refs}{}
\begin{CSLReferences}{1}{0}
\leavevmode\hypertarget{ref-Agosta2010HowSpe}{}%
Agosta, S. J. et al. 2010. How specialists can be generalists: Resolving
the {``parasite paradox''} and implications for emerging infectious
disease. - Zoologia (Curitiba) 27: 151--162.

\leavevmode\hypertarget{ref-Albery2020PreGlo}{}%
Albery, G. F. et al. 2020. Predicting the global mammalian viral sharing
network using phylogeography. - Nature Communications 11: 2260.

\leavevmode\hypertarget{ref-Allen2017GloHot}{}%
Allen, T. et al. 2017. Global hotspots and correlates of emerging
zoonotic diseases. - Nature Communications in press.

\leavevmode\hypertarget{ref-Alves2018GeoVar}{}%
Alves, D. M. C. C. et al. 2018. Geographic variation in the relationship
between large-scale environmental determinants and bat species richness.
- Basic and Applied Ecology 27: 1--8.

\leavevmode\hypertarget{ref-Anthony2017GloPat}{}%
Anthony, S. J. et al. 2017. Global patterns in coronavirus diversity. -
Virus Evolution in press.

\leavevmode\hypertarget{ref-Banerjee2020NovIns}{}%
Banerjee, A. et al. 2020. Novel Insights Into Immune Systems of Bats. -
Frontiers in Immunology 11: 26.

\leavevmode\hypertarget{ref-Becker2022OptPre}{}%
Becker, D. J. et al. 2022. Optimising predictive models to prioritise
viral discovery in zoonotic reservoirs. - The Lancet Microbe in press.

\leavevmode\hypertarget{ref-Calisher2006BatImp}{}%
Calisher, C. H. et al. 2006. Bats: Important Reservoir Hosts of Emerging
Viruses. - Clinical Microbiology Reviews 19: 531--545.

\leavevmode\hypertarget{ref-Carlson2022CliCha}{}%
Carlson, C. J. et al. 2022. Climate change increases cross-species viral
transmission risk. - Nature: 1--1.

\leavevmode\hypertarget{ref-Cavender-Bares2009MerCom}{}%
Cavender-Bares, J. et al. 2009. The merging of community ecology and
phylogenetic biology. - Ecol. Lett. 12: 693--715.

\leavevmode\hypertarget{ref-Cohen2022SamStr}{}%
Cohen, L. E. et al. 2022. Sampling strategies and pre-pandemic
surveillance gaps for bat coronaviruses.: 2022.06.15.496296.

\leavevmode\hypertarget{ref-Dansereau2022EvaEco}{}%
Dansereau, G. et al. 2022. Evaluating ecological uniqueness over broad
spatial extents using species distribution modelling. - Oikos n/a:
e09063.

\leavevmode\hypertarget{ref-Deshpande2022ForFru}{}%
Deshpande, K. et al. 2022. Forbidden fruits? Ecosystem services from
seed dispersal by fruit bats in the context of latent zoonotic risk. -
Oikos (Copenhagen, Denmark): oik.08359.

\leavevmode\hypertarget{ref-Drexler2014EcoEvo}{}%
Drexler, J. F. et al. 2014. Ecology, evolution and classification of bat
coronaviruses in the aftermath of SARS. - Antiviral Research 101:
45--56.

\leavevmode\hypertarget{ref-Faith1992ConEva}{}%
Faith, D. P. 1992. Conservation evaluation and phylogenetic diversity. -
Biological Conservation 61: 1--10.

\leavevmode\hypertarget{ref-Gomulkiewicz2000HotSpo}{}%
Gomulkiewicz, R. et al. 2000. Hot Spots, Cold Spots, and the Geographic
Mosaic Theory of Coevolution. - The American Naturalist 156: 156--174.

\leavevmode\hypertarget{ref-Gomulkiewicz2007DosDon}{}%
Gomulkiewicz, R. et al. 2007. Dos and don'ts of testing the geographic
mosaic theory of coevolution. - Heredity 98: 249--258.

\leavevmode\hypertarget{ref-Gonsalves2013MosInf}{}%
Gonsalves, L. et al. 2013a. Do mosquitoes influence bat activity in
coastal habitats? - Wildlife Research 40: 10--24.

\leavevmode\hypertarget{ref-Gonsalves2013MosCon}{}%
Gonsalves, L. et al. 2013b. Mosquito Consumption by Insectivorous Bats:
Does Size Matter? - PLOS ONE 8: e77183.

\leavevmode\hypertarget{ref-Gorbunova2020WorGoe}{}%
Gorbunova, V. et al. 2020. The World Goes Bats: Living Longer and
Tolerating Viruses. - Cell Metabolism 32: 31--43.

\leavevmode\hypertarget{ref-Gryseels2017WheVir}{}%
Gryseels, S. et al. 2017. When Viruses Don't Go Viral: The Importance of
Host Phylogeographic Structure in the Spatial Spread of Arenaviruses (JH
Kuhn, Ed.). - PLOS Pathogens 13: e1006073.

\leavevmode\hypertarget{ref-Hazarie2021IntPop}{}%
Hazarie, S. et al. 2021. Interplay between population density and
mobility in determining the spread of epidemics in cities. -
Communications Physics 4: 191.

\leavevmode\hypertarget{ref-Hoang2018UfbImp}{}%
Hoang, D. T. et al. 2018. UFBoot2: Improving the Ultrafast Bootstrap
Approximation. - Molecular Biology and Evolution 35: 518--522.

\leavevmode\hypertarget{ref-Irving2021LesHos}{}%
Irving, A. T. et al. 2021. Lessons from the host defences of bats, a
unique viral reservoir. - Nature 589: 363--370.

\leavevmode\hypertarget{ref-Isaac2007MamEdg}{}%
Isaac, N. J. B. et al. 2007. Mammals on the EDGE: Conservation
Priorities Based on Threat and Phylogeny. - PLOS ONE 2: e296.

\leavevmode\hypertarget{ref-IUCN2021IucRed}{}%
IUCN 2021. The IUCN Red List of Threatened Species.

\leavevmode\hypertarget{ref-Janzen1980WheIt}{}%
Janzen, D. H. 1980. When is it Coevolution? - Evolution 34: 611--612.

\leavevmode\hypertarget{ref-Jebb2020SixRef}{}%
Jebb, D. et al. 2020. Six reference-quality genomes reveal evolution of
bat adaptations. - Nature 583: 578--584.

\leavevmode\hypertarget{ref-Johnson2020GloShi}{}%
Johnson, C. K. et al. 2020. Global shifts in mammalian population trends
reveal key predictors of virus spillover risk. - Proceedings of the
Royal Society B: Biological Sciences 287: 20192736.

\leavevmode\hypertarget{ref-Kalyaanamoorthy2017ModFas}{}%
Kalyaanamoorthy, S. et al. 2017. ModelFinder: Fast model selection for
accurate phylogenetic estimates. - Nature Methods 14: 587--589.

\leavevmode\hypertarget{ref-Kasso2013EcoEco}{}%
Kasso, M. and Balakrishnan, M. 2013. Ecological and Economic Importance
of Bats (Order Chiroptera). - ISRN Biodiversity 2013: e187415.

\leavevmode\hypertarget{ref-Katoh2013MafMul}{}%
Katoh, K. and Standley, D. M. 2013. MAFFT Multiple Sequence Alignment
Software Version 7: Improvements in Performance and Usability. -
Molecular Biology and Evolution 30: 772--780.

\leavevmode\hypertarget{ref-Keke2010StuSki}{}%
Keke, S. et al. 2010. Study on skin color image segmentation used by
Fuzzy-c-means arithmetic. - 2010 Seventh International Conference on
Fuzzy Systems and Knowledge Discovery 2: 612--615.

\leavevmode\hypertarget{ref-Kettenburg2022FulGen}{}%
Kettenburg, G. et al. 2022. Full Genome Nobecovirus Sequences From
Malagasy Fruit Bats Define a Unique Evolutionary History for This
Coronavirus Clade. - Frontiers in Public Health in press.

\leavevmode\hypertarget{ref-Kreft2007GloPat}{}%
Kreft, H. and Jetz, W. 2007. Global patterns and determinants of
vascular plant diversity. - Proceedings of the National Academy of
Sciences 104: 5925--5930.

\leavevmode\hypertarget{ref-Kreft2010FraDel}{}%
Kreft, H. and Jetz, W. 2010. A framework for delineating biogeographical
regions based on species distributions. - Journal of Biogeography 37:
2029--2053.

\leavevmode\hypertarget{ref-Latinne2020OriCro}{}%
Latinne, A. et al. 2020. Origin and cross-species transmission of bat
coronaviruses in China. - bioRxiv: The Preprint Server for Biology:
2020.05.31.116061.

\leavevmode\hypertarget{ref-Lednicky2021IsoNov}{}%
Lednicky, J. A. et al. 2021. Isolation of a Novel Recombinant Canine
Coronavirus from a Visitor to Haiti: Further Evidence of Transmission of
Coronaviruses of Zoonotic Origin to Humans. - Clinical Infectious
Diseases: An Official Publication of the Infectious Diseases Society of
America: ciab924.

\leavevmode\hypertarget{ref-Legendre2013BetDiv}{}%
Legendre, P. and De Cáceres, M. 2013. Beta diversity as the variance of
community data: Dissimilarity coefficients and partitioning (H Morlon,
Ed.). - Ecology Letters 16: 951--963.

\leavevmode\hypertarget{ref-Legendre2019SpaTem}{}%
Legendre, P. and Condit, R. 2019. Spatial and temporal analysis of beta
diversity in the Barro Colorado Island forest dynamics plot, Panama. -
Forest Ecosystems 6: 7.

\leavevmode\hypertarget{ref-Letko2020BatVir}{}%
Letko, M. et al. 2020. Bat-borne virus diversity, spillover and
emergence. - Nature Reviews Microbiology 18: 461--471.

\leavevmode\hypertarget{ref-Mello2011MisPar}{}%
Mello, M. A. R. et al. 2011. The Missing Part of Seed Dispersal
Networks: Structure and Robustness of Bat-Fruit Interactions. - PLOS ONE
6: e17395.

>>>>

\noindent {\LARGE \textbf{\textsf{The coevolutionary mosaic of
betacoronavirus emergence risk}}}

@@

We downloaded the data on bats hosts of \emph{Betacoronavirus} assembled
by\textsuperscript{20} from
\texttt{https://www.viralemergence.org/betacov} on Apr.~2022, and
filtered it to ``known'' hosts (established before the emergence of
SARS-CoV-2) and ``novel'' hosts (confirmed through sampling and
competence assays since the initial data collection). The original
database was assembled by a combination of data mining and literature
surveys, including automated alerts on the ``bats'' and ``coronavirus''
keywords to identify novel empirical evidence of bats-betacoronaviruses
associations; this yielded a total of 126 known hosts, 47 of which were
novel hosts.

@@

We downloaded the rangemap of every current bat species that was
classified as an empirically documented host of \emph{Betacoronavirus}
from the previous step, according to recent IUCN
data.\textsuperscript{54} The range maps were subsequently rasterized
using the \texttt{rasterize} function from
\texttt{GDAL}\textsuperscript{55} at a resolution of approximately
100kmx100km. For every pixel in the resulting raster where at least one
bat host of \emph{Betacoronavirus} was present, we extract the species
pool (list of all known bat hosts), which was used to calculate the
following risk assessment components: bat phylogenetic diversity, bat
compositional uniqueness, and predicted viral sharing risk.

@@

For every pixel, we measured Faith's Phylogenetic
Diversity\textsuperscript{56} based on a recent synthetic tree with
robust time calibration, covering about 6000 mammalian
species.\textsuperscript{57} Faith's PD measures the sum of unique
branches from an arbitrary root to a set of tips, and comparatively
larger values indicate a more phylogenetic diverse species pool. We
measured phylogenetic diversity starting from the root of the entire
tree (and not from Chiroptera); this bears no consequences on the
resulting values, since all branches leading up to Chiroptera are only
counted one per species pool, and (as we explain when describing the
assembly of the composite risk map), all individual risk components are
ranged in {[}0,1{]}. This measure incorporates a richness component,
which we chose not to correct for; the interpretation of the
phylogenetic diversity is therefore a weighted species richness, that
accounts for phylogenetic over/under-dispersal in some places.

@@

For every species pool, we measured its Local Contribution to
Beta-Diversity;\textsuperscript{58} LCBD works from a species-data
matrix (traditionally noted as \(\mathbf{Y}\)), where species are rows
and sites are columns, and a value of 1 indicates occurrence. We
extracted the Y matrix assuming that every pixel represents a unique
location, and following best practices\textsuperscript{59} transformed
it using Hellinger's distance to account for unequal bat richness at
different pixels. The correction of raw community data is particularly
important for two reasons: first, it prevents the artifact of richer
sites having higher importance; second, it removes the effect of overall
species richness, which is already incorporated in the phylogenetic
diversity component. High values of LCBD indicate that the pixel has a
community that is on average more dissimilar in species composition than
what is expected knowing the entire matrix, i.e.~a more unique
community. Recent results by\textsuperscript{60} shows that LCBD
measures are robust with regards to spatial scale, and are therefore
applicable at the global scale.

@@

For all bat hosts of \emph{Betacoronavirus}, we extracted their
predicted viral sharing network, generated from a previously published
generalized additive mixed model of virus sharing by a tensor function
of phylogenetic distance and geographic range
overlap.\textsuperscript{61} This network stores pairwise values of
viral community similarity. To project viral sharing values into a
single value for every pixel, we averaged the pairwise scores. High
values of the average sharing propensity means that this specific extant
bat assemblage is likely to be proficient at exchanging viruses.

@@

To visualize the aggregated risk at the global scale, we combine the
three individual risk components (phylogenetic diversity, compositional
uniqueness, and viral sharing) using an additive color
model.\textsuperscript{62} In this approach, every risk component gets
assigned a component in the RGB color model (phylogenetic diversity is
green, compositional uniqueness is red, and viral sharing is blue). In
order to achieve a valid RGB measure, all components are re-scaled to
the {[}0,1{]} interval, so that a pixel with no sharing, no phylogenetic
diversity, and no compositional uniqueness is black, and a pixel with
maximal values for each is white. This additive model conveys both the
intensity of the overall risk, but also the nature of the risk as colors
diverge towards combinations of values for three risk components. Out of
the possible combinations, the most risky in terms or rapid
diversification and spillover potential is high phylogenetic diversity
and low viral sharing,\textsuperscript{63,64} in that this allows
multiple independent host-virus coevolutionary dynamics to take place in
the same location. In the colorimetric space, this correspond to yellow
-- because the HSV space is more amenable to calculations for feature
extraction,\textsuperscript{see \emph{e.g.} 65} we measured the risk
level by calculating the angular distance of the hue of each pixel to a
reference value of 60, and weighted this risk level by the value
component. Specifically, given a pixel with colorimetric coordinates
\((h,s,v)\), its ranged weighted risk value is

@@

To next represent phylogeography of betacoronaviruses in bats, we
aggregated and analyzed betacoronavirus sequence data. We used the
following query to pull all \emph{Betacoronavirus} sequence data from
the GenBank Nucleotide database except SARS-CoV-2;
(``Betacoronavirus''{[}Organism{]} OR betacoronavirus{[}All Fields{]})
NOT (``Severe acute respiratory syndrome coronavirus 2''{[}Organism{]}
OR sars-cov-2{[}All Fields{]}). We added a single representative
sequence for SARS-CoV-2 and manually curated to remove sequences without
the RNA-dependent RNA polymerase (RdRp) sequence or that contained words
indicating recombinant or laboratory strains including ``patent'',
``mutant'', ``GFP'', and ``recombinant''. We filtered over-represented
taxa including betacoronavirus 1, hCoV-OC43, Middle East respiratory
syndrome coronavirus, Murine hepatitis virus, and hCoV-HKU1. Curated
betacoronavirus RdRp sequences were then aligned using MAFFT v 1.4.0
{[}\textsuperscript{66}; Algorithm FFT-NS-2, Scoring matrix 200PAM /
k=2, gap open penalty 1.53m offset value 0.123{]} and a maximum
likelihood tree reconstructed in IQ-TREE v 1.6.12\textsuperscript{67}
with ModelFinder\textsuperscript{68} ultrafast bootstrap
approximation\textsuperscript{69} with a general time reversible model
with empirical base frequencies and the 5-discrete-rate-category
FreeRaye model of nucleotide substitution (GTR+F+R5).

We first tested the hypothesis that hotspots of viral diversification
would track hotspots of bat diversification. To do so, we plotted the
number of known bat hosts (specifically only those included in the
phylogeny, so there was a 1:1 correspondence between data sources)
against the ``mean evolutionary distinctiveness'' of the associated
viruses. To calculate this, we derived the fair proportions evolutionary
distinctiveness\textsuperscript{70} for each of the viruses in the tree,
then averaged these at the bat species level, projected these values
onto their geographic distributions, and averaged across every bat found
in a given pixel. As such, this can be thought of as a map of the mean
evolutionary distinctiveness of the known viral community believed to be
associated with a particular subset of bats present.

@@

Subsequently, we tested the hypothesis that the biogeography of bat
betacoronaviruses should track the biogeography of their hosts. To test
this idea, we loosely adapted a method from,\textsuperscript{71,72} who
proposed a phylogenetic method for the delineation of animal
biogeographic regions. In their original method, a distance matrix -
where each row or column represents a geographic raster's grid cell, and
the dissimilarity values are the ``beta diversity similarity'' of their
community assemble - undergoes non-metric multidimensional scaling
(NMDS); the first two axes of the NMDS are projected geographically
using a four-color bivariate map. Here, we build on this idea with an
entirely novel methodology. First, we measure the phylogenetic distance
between the different viruses in the betacoronaviruses tree by using the
cophenetic function in \texttt{ape};\textsuperscript{73} subsequently,
we take a principal components analysis of that distance matrix (readily
interchangeable for NMDS in this case) to project the viral tree into an
n-dimensional space. We then take the first two principal components
and, as with the evolutionary distinctiveness analysis, aggregated these
to a mean host value and projected them using a four-color bivariate
map.

@@

\hypertarget{refs}{}
\begin{CSLReferences}{0}{0}
\leavevmode\hypertarget{ref-Plowright2017PatZoo}{}%
\CSLLeftMargin{1. }
\CSLRightInline{Plowright, R. K. \emph{et al.} Pathways to zoonotic
spillover. \emph{Nature Reviews Microbiology} \textbf{15}, 502--510
(2017).}

\leavevmode\hypertarget{ref-Anthony2017GloPat}{}%
\CSLLeftMargin{2. }
\CSLRightInline{Anthony, S. J. \emph{et al.} Global patterns in
coronavirus diversity. \emph{Virus Evolution} \textbf{3}, (2017).}

\leavevmode\hypertarget{ref-Ruiz-Aravena2022EcoEvo}{}%
\CSLLeftMargin{3. }
\CSLRightInline{Ruiz-Aravena, M. \emph{et al.} Ecology, evolution and
spillover of coronaviruses from bats. \emph{Nature Reviews Microbiology}
\textbf{20}, 299--314 (2022).}

\leavevmode\hypertarget{ref-Agosta2010HowSpe}{}%
\CSLLeftMargin{4. }
\CSLRightInline{Agosta, S. J., Janz, N. \& Brooks, D. R. How specialists
can be generalists: Resolving the {`parasite paradox'} and implications
for emerging infectious disease. \emph{Zoologia (Curitiba)} \textbf{27},
151--162 (2010).}

\leavevmode\hypertarget{ref-Power2004PatSpi}{}%
\CSLLeftMargin{5. }
\CSLRightInline{Power, A. G. \& Mitchell, C. E. Pathogen Spillover in
Disease Epidemics. \emph{The American Naturalist} \textbf{164}, S79--S89
(2004).}

\leavevmode\hypertarget{ref-Thompson2005GeoMos}{}%
\CSLLeftMargin{6. }
\CSLRightInline{Thompson, J. N. \emph{The Geographic Mosaic of
Coevolution}. (University Of Chicago Press, 2005).}

\leavevmode\hypertarget{ref-Thompson1994CoePro}{}%
\CSLLeftMargin{7. }
\CSLRightInline{Thompson, J. N. \emph{The Coevolutionary Process}.
(University of Chicago Press, 1994).}

\leavevmode\hypertarget{ref-Janzen1980WheIt}{}%
\CSLLeftMargin{8. }
\CSLRightInline{Janzen, D. H. When is it Coevolution? \emph{Evolution}
\textbf{34}, 611--612 (1980).}

\leavevmode\hypertarget{ref-Price2002MacThe}{}%
\CSLLeftMargin{9. }
\CSLRightInline{Price, P. W. \emph{Macroevolutionary Theory on
Macroecological Patterns}. (Cambridge University Press, 2002).
doi:\href{https://doi.org/10.1017/CBO9780511615030}{10.1017/CBO9780511615030}.}

\leavevmode\hypertarget{ref-Gomulkiewicz2007DosDon}{}%
\CSLLeftMargin{10. }
\CSLRightInline{Gomulkiewicz, R. \emph{et al.} Dos and don'ts of testing
the geographic mosaic theory of coevolution. \emph{Heredity}
\textbf{98}, 249--258 (2007).}

\leavevmode\hypertarget{ref-Thrall2007CoeSym}{}%
\CSLLeftMargin{11. }
\CSLRightInline{Thrall, P. H., Hochberg, M. E., Burdon, J. J. \& Bever,
J. D. Coevolution of symbiotic mutualists and parasites in a community
context. \emph{Trends in Ecology \& Evolution} \textbf{22}, 120--126
(2007).}

\leavevmode\hypertarget{ref-Drexler2014EcoEvo}{}%
\CSLLeftMargin{12. }
\CSLRightInline{Drexler, J. F., Corman, V. M. \& Drosten, C. Ecology,
evolution and classification of bat coronaviruses in the aftermath of
SARS. \emph{Antiviral Research} \textbf{101}, 45--56 (2014).}

\leavevmode\hypertarget{ref-Calisher2006BatImp}{}%
\CSLLeftMargin{13. }
\CSLRightInline{Calisher, C. H., Childs, J. E., Field, H. E., Holmes, K.
V. \& Schountz, T. Bats: Important Reservoir Hosts of Emerging Viruses.
\emph{Clinical Microbiology Reviews} \textbf{19}, 531--545 (2006).}

\leavevmode\hypertarget{ref-Moratelli2015BatZoo}{}%
\CSLLeftMargin{14. }
\CSLRightInline{Moratelli, R. \& Calisher, C. H. Bats and zoonotic
viruses: Can we confidently link bats with emerging deadly viruses?
\emph{Memórias do Instituto Oswaldo Cruz} \textbf{110}, 1--22 (2015).}

\leavevmode\hypertarget{ref-Mollentze2020VirZoo}{}%
\CSLLeftMargin{15. }
\CSLRightInline{Mollentze, N. \& Streicker, D. G. Viral zoonotic risk is
homogenous among taxonomic orders of mammalian and avian reservoir
hosts. \emph{Proceedings of the National Academy of Sciences}
\textbf{117}, 9423--9430 (2020).}

\leavevmode\hypertarget{ref-Carlson2022CliCha}{}%
\CSLLeftMargin{16. }
\CSLRightInline{Carlson, C. J. \emph{et al.} Climate change increases
cross-species viral transmission risk. \emph{Nature} 1--1 (2022)
doi:\href{https://doi.org/10.1038/s41586-022-04788-w}{10.1038/s41586-022-04788-w}.}

\leavevmode\hypertarget{ref-Springer2013PhyBat}{}%
\CSLLeftMargin{17. }
\CSLRightInline{Springer, M. S. Phylogenetics: Bats United, Microbats
Divided. \emph{Current Biology} \textbf{23}, R999--R1001 (2013).}

\leavevmode\hypertarget{ref-Teeling2005MolPhy}{}%
\CSLLeftMargin{18. }
\CSLRightInline{Teeling, E. C. \emph{et al.} A Molecular Phylogeny for
Bats Illuminates Biogeography and the Fossil Record. \emph{Science (New
York, N.Y.)} \textbf{307}, 580--584 (2005).}

\leavevmode\hypertarget{ref-Villalobos2010DivFie}{}%
\CSLLeftMargin{19. }
\CSLRightInline{Villalobos, F. \& Arita, H. T. The diversity field of
New World leaf-nosed bats (Phyllostomidae). \emph{Global Ecology and
Biogeography} \textbf{19}, 200--211 (2010).}

\leavevmode\hypertarget{ref-Becker2022OptPre}{}%
\CSLLeftMargin{20. }
\CSLRightInline{Becker, D. J. \emph{et al.} Optimising predictive models
to prioritise viral discovery in zoonotic reservoirs. \emph{The Lancet
Microbe} (2022)
doi:\href{https://doi.org/10.1016/S2666-5247(21)00245-7}{10.1016/S2666-5247(21)00245-7}.}

\leavevmode\hypertarget{ref-Allen2017GloHot}{}%
\CSLLeftMargin{21. }
\CSLRightInline{Allen, T. \emph{et al.} Global hotspots and correlates
of emerging zoonotic diseases. \emph{Nature Communications} \textbf{8},
(2017).}

\leavevmode\hypertarget{ref-Olival2017HosVir}{}%
\CSLLeftMargin{22. }
\CSLRightInline{Olival, K. J. \emph{et al.} Host and viral traits
predict zoonotic spillover from mammals. \emph{Nature} \textbf{546},
646--650 (2017).}

\leavevmode\hypertarget{ref-VanBrussel2022ZooDis}{}%
\CSLLeftMargin{23. }
\CSLRightInline{Van Brussel, K. \& Holmes, E. C. Zoonotic disease and
virome diversity in bats. \emph{Current Opinion in Virology}
\textbf{52}, 192--202 (2022).}

\leavevmode\hypertarget{ref-Letko2020BatVir}{}%
\CSLLeftMargin{24. }
\CSLRightInline{Letko, M., Seifert, S. N., Olival, K. J., Plowright, R.
K. \& Munster, V. J. Bat-borne virus diversity, spillover and emergence.
\emph{Nature Reviews Microbiology} \textbf{18}, 461--471 (2020).}

\leavevmode\hypertarget{ref-Banerjee2020NovIns}{}%
\CSLLeftMargin{25. }
\CSLRightInline{Banerjee, A. \emph{et al.} Novel Insights Into Immune
Systems of Bats. \emph{Frontiers in Immunology} \textbf{11}, 26 (2020).}

\leavevmode\hypertarget{ref-Gorbunova2020WorGoe}{}%
\CSLLeftMargin{26. }
\CSLRightInline{Gorbunova, V., Seluanov, A. \& Kennedy, B. K. The World
Goes Bats: Living Longer and Tolerating Viruses. \emph{Cell Metabolism}
\textbf{32}, 31--43 (2020).}

\leavevmode\hypertarget{ref-Irving2021LesHos}{}%
\CSLLeftMargin{27. }
\CSLRightInline{Irving, A. T., Ahn, M., Goh, G., Anderson, D. E. \&
Wang, L.-F. Lessons from the host defences of bats, a unique viral
reservoir. \emph{Nature} \textbf{589}, 363--370 (2021).}

\leavevmode\hypertarget{ref-Jebb2020SixRef}{}%
\CSLLeftMargin{28. }
\CSLRightInline{Jebb, D. \emph{et al.} Six reference-quality genomes
reveal evolution of bat adaptations. \emph{Nature} \textbf{583},
578--584 (2020).}

\leavevmode\hypertarget{ref-MorenoSantillan2021LarGen}{}%
\CSLLeftMargin{29. }
\CSLRightInline{Moreno Santillán, D. D. \emph{et al.} Large-scale genome
sampling reveals unique immunity and metabolic adaptations in bats.
\emph{Molecular Ecology} mec.16027 (2021)
doi:\href{https://doi.org/10.1111/mec.16027}{10.1111/mec.16027}.}
